---
title: "Determinate sentence bands"
author: "Brian G Burton"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Determinate sentence bands}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

<!---
```{r setup}
#library(prisonsreadyreckoner)
```
--->


# Definitions{#sentence-bands}

Modelling of determinate sentence prisoners is on the basis of four sentence
length bands. Relevant levers are defined in these terms. Bands are:

1. Sentence band 1: Prisoners serving a sentence shorter than six months.

2. Sentence band 2: Prisoners serving a sentence of length of at least six months
and shorter than twelve months.

3. Sentence band 3: Prisoners serving a sentence of length of at least twelve
months and shorter than four years.

4. Sentence band 4: Prisoners serving a sentence of length of at least four
years. Extended Determinate Sentence (EDS) prisoners are included within this
band.
