---
title: "Updating prisonsreadyreckoner"
author: "Alex Dickinson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Updating prisonsreadyreckoner}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = TRUE,
  comment = "#>"
)
```

```{css, echo=FALSE}
.html-widget {
    height: 100% !important;
    width: 100% !important;
}
```

<!---
```{r setup}
#library(mojmodel)
```
--->


Assuming that underlying modelling assumptions do not change, and changes are due simply to updates to Central forecast files.

To update `prisonsreadyreckoner`, follow these steps:

1. Obtain updated versions of files that are . Some of these files are simply copies of files supplied directly by modelling teams. Other files are created using the package `prisonsreadyreckonerupdater`.
 
* **Upstream and Courts Modelling team**. Team leader for current version: Ceri Cooper.
  +	**File:** `police_charges_cc_files`. Analyst for current version: David Verry.
  + **File:** `police_charges_mc_file`. Analyst for current version: Katie Mahon.
  + **File:** `police_charges_cc_route_file`. Analyst for current version: Chun-yee Cheng.
  + **File:** `cc_output_file`. Analyst for current version: Chun-yee Cheng.
  + **File:** `cc_capacity_file`. Analyst for current version: Chun-yee Cheng.
* **Prison Demand Modelling team**. Team leader for current version: Charlotte Wallace.
  + **File:** `prison_inflows_file`. Analyst for current version: Charlotte Wallace.
  + **File:** `recall_file`. Analyst for current version: Charlotte Wallace.
  + **File:** `licence_profiles_file`. Analyst for current version: Jordan Carroll.
  + **Numerical parameters:** `remand_rates`. Analyst for current version: Jordan Carroll.

prisonsreadyreckoner also relies on files that are created using the package `prisonsreadyreckonerupdater`. To update these files, follow these steps:

*	**File:** `sentencing_rates_file`.
  + Obtain updated version of **file** `path_rdo_data`. Maintained by Prison Demand Modelling team. Current version provided by Claire O’Neill.
  + Obtain updated version of **file** `path_cc_disposals`. Maintained by Prison Demand Modelling team. Current version provided by Claire O’Neill.
  + Run `prisonsreadyreckoner` function `update_reception_rates_old()`. This function uses the old version of fullsample.
* **File:** `profiles_file`.
  + Obtain updated version of **file** `path_flows` – Maintained by Prison Demand Modelling team. Current version provided by Charlotte Wallace.
  + Run `prisonsreadyreckoner` function `update_determinate_profiles_old()`. This function uses the old version of fullsample.
* **File:** `gender_splits_file`. Run `prisonsreadyreckoner` function `update_gender_splits_old()`. This function uses the old version of fullsample.

The following files are unchanged:

* **File:** `ringfenced_lookup_file`.

2)	Run prisonsreadyreckonerupdater
3)	Decide app parameters

Subsidiary details for use in app:
1)	Historical data
2)	Population file
3)	Capacity file




<!-- Cross-vignette references -->
[parameters]: ./parameters.html#parameters "Parameters"
[sentence-bands]: ./sentence-bands.html#sentence-bands "Sentence bands"







