---
title: "User guide"
author: "Alex Dickinson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{User guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = TRUE,
  comment = "#>"
)
```

```{css, echo=FALSE}
.html-widget {
    height: 100% !important;
    width: 100% !important;
}
```

<!---
```{r setup}
#library(mojmodel)
```
--->


The model `prisonsreadyreckoner` generates projections of how the prison
population will change in future. These projections are influenced by five
**Levers**, which may all be set by adjusting the values of the relevant
[parameters][parameters].

<div>
```{r, echo=FALSE}
mojflowchart::flowchart("prisonsreadyreckoner-flow-chart.mmd")
```
</div>

## Vary number of offenders entering prison

The user can vary the number of offenders entering prison using a combination of
**Lever 1**, **Lever 2** and **Lever 3**. **Lever 1** and **Lever 2** capture
the effects of changes to the number of court cases and to Crown Court capacity,
whereas **Lever 3** simply adds a specified number of offenders beginning
determinate prison sentences. Although all three levers can be varied
simultaneously, to gain an understanding of the effect of changes to the
criminal justice system we strongly recommend that you pull either **Lever 1**
and **Lever 2** or **Lever 3**.


### **Lever 1**: Vary number of court cases

The user can load scenarios that capture changes to the number of cases received
by the magistrates' courts and the Crown Courts. The user is free to define
these scenarios. Scenarios could reflect, for instance, increases in received
cases due to increases in the number of charges brought by police.

See functions `load_police_charges_cc_data_list()` and
`load_police_charges_mc_data()`.


### **Lever 2**: Vary number of court sitting days {#lever-vary-number-of-court-sitting-days}

The user can add an additional number of court sitting days per year, beginning
on a specified **Impact date**.

See function `add_cc_sitting_days()`.


### **Lever 3**: Add additional determinate sentence prisoners {#lever-add-additional-determinate-sentence-prisoners}

The user can specify an additional number of offenders who will begin
determinate prison sentences each month. Different changes can be specified for
different [Sentence bands][sentence-bands], and changes will come into effect on
a specified **Impact date**. Note that this lever is completely independent from
**Lever 1** and **Lever 2**, and provides a way to simply add an arbitrary
number of determinate sentence prisoners.

See function `add_inflows_det_delta_lever()`.


## Vary time served in prison

### **Lever 4**: Vary time served in prison by determinate sentence prisoners {#lever-vary-time-served-in-prison-by-determinate-sentence-prisoners}

The user can scale the time served in prison by determinate sentence prisoners
by a specified factor. Different factors can be specified for each [Sentence
band][sentence-bands], and will affect offenders who enter prison after a
specified **Impact date**. Note that this lever adjusts only the time served
during a prisoner’s sentence, and does not adjust the time spent on remand or in
recall.

See function `stretch_profiles()`.


## Vary number of licensees recalled to prison

### **Lever 5**: Vary percentage of determinate sentence offenders recalled to prison {#lever-vary-percentage-of-offenders-recalled-to-prison}

The user can change the percentage of determinate sentence offenders who are
released on licence and subsequently recalled to prison (per month). Different
percentages can be specified for each [Sentence band][sentence-bands], and changes will come into
effect on a specified **Impact date**. Note that this lever affects only offenders
who enter prison in the forecast period and who are subsequently released on licence.


## Vary time served by licensees recalled to prison

### **Lever 6**: Scale time served in prison by recalled licensees

The user can scale the time served in prison by recalled licensees by a selected
factor. For instance, selecting a factor of 2 will double the time served.
Different factors can be specified for each [Sentence band][sentence-bands] and
will affect licensees who are recalled after a specified **Impact date**.

See function `stretch_recall_time_lever()`.


<!-- Cross-vignette references -->
[parameters]: ./parameters.html#parameters "Parameters"
[sentence-bands]: ./sentence-bands.html#sentence-bands "Sentence bands"







